### 2013-03-28 doc/sosp13/evaluation.tex
commit to merge

\subsection{Basic \Grappa Mechanisms}

\paragraph{User-level context switching.}
As discussed earlier, fast context switching is at the heart of \Grappa's
latency tolerance abilities. We assess context switch overheads using a simple
microbenchmark that has a configurable number of workers, where each worker 
just increments values in a large array.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/context_switch_time.pdf}
%      \includegraphics[width=0.5\textwidth]{figs/context_switch_bw.pdf}
    \end{center}
    \caption{Average context switch time with 1 active, with and without prefetching.}
    \label{fig:context-switch-exp}
\end{figure}

Figure~\ref{fig:context-switch-exp} shows the average context switch time as
the number of workers grow. There are several important takeaway points form
these data. Context switch time for the number of workers we use currently
($~$1K) is on the order of 50 ns; and prefetching has significant impact on
context switch time. And ontext switch time does not grow significantly with
the number of workers --- even at close to one half million workers context
switch is around 75ns.

@@

In summary, our context switch engine is able to efficiently sustain very high concurrency and we will show later, the amount of concurrency sustained is sufficient the latencies \Grappa needs to hide.

\paragraph{Global memory and communication.} We measure the performance of
\Grappa's global memory and communication layers using a faithful
implementation of the giga updates per second (GUPS) benchmark.
Read-modify-write updates are dispatched at random to a large array. This
benchmark stresses the communication layer of \Grappa separately from the
scheduler, because only a single worker is used per system node. Figure~\ref{fig:grappa-gups} shows that \Grappa is able to sustain on the order of billions of updates per second, which is enough to \TODO{...}.

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-gups.pdf}
    \end{center}
    \caption{GUPS (giga updates per second) for \Grappa as the number of nodes grows.}
    \label{fig:grappa-gups}
\end{figure}

@@

\begin{figure}[ht]
    \begin{center}
%      \includegraphics[width=0.5\textwidth]{figs/grappa-uts.pdf}
    \end{center}
    \caption{Vertices per second in UTS on \Grappa as the number of nodes grows.}
    \label{fig:grappa-uts}
\end{figure}
\TODO{UTS: story here is T1 class is bushy trees similar to properties
    of real world networks. Larger T1XXL (4.5 B vertices) provides
    more parallelism for continued scaling up. T3 class is deep skinny
    trees. Grappa can do no better than the straightforward UPC impl: there very little
    parallelism and the critical path is long}

@@

In order to ground \Grappa's performance, we compare it with XMT running BFS,
PageRank, IntSort, GUPS and UTS. Since XMT is a different hardware platform,
we also compare \Grappa with hand-tuned MPI versions of BFS and GUPS running
on the same hardware. Finally, we also compare it with UTS written for UPC. We
run all experiments with 64 nodes.

>>>>

\subsection{Basic \Grappa Mechanisms}
\label{eval:basic}

\paragraph{User-level context switching.}

As discussed earlier, fast context switching is at the heart of \Grappa's
latency tolerance abilities. We assess context switch overheads using a simple
microbenchmark that has a configurable number of workers, where each worker 
just increments values in a large array.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/context_switch_time.pdf}
    \end{center}
    \caption{Average context switch time with and without prefetching.}
    \label{fig:context-switch-exp}
\end{figure}

Figure~\ref{fig:context-switch-exp} shows the average context switch time as
the number of workers grow. There are two important takeaway points from these
data. Context switch time for the number of workers we use ($~$1K workers) is
on the order of 50 ns; and prefetching has significant impact on context
switch time. And context switch time does not grow significantly with the
number of workers --- even with one-half million workers, context switch
is around 75ns.

@@

In summary, our context switch engine is able to efficiently sustain very high
concurrency and as we will show later, the amount of concurrency sustained is
sufficient the latencies \Grappa needs to hide.

\paragraph{Global memory and communication.} We measure the performance of
\Grappa's global memory and communication layers using a faithful
implementation of the giga updates per second (GUPS) benchmark.
Read-modify-write updates are dispatched at random to a large array. This
benchmark stresses the communication layer of \Grappa separately from the
scheduler, because only a single worker is used per system node.
Figure~\ref{fig:grappa-gups} shows that \Grappa is able to sustain well over a
billion updates per second with 64 nodes. This compares very favorably to
published results~\footnote{\url{http://icl.cs.utk.edu/hpcc/hpcc_results.cgi}}
for other high-end HPC systems.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{results/gups/gups.pdf}
    \end{center}
    \caption{GUPS (giga updates per second) for \Grappa as the number of nodes grows.}
    \label{fig:grappa-gups}
\end{figure}

@@

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_scale.pdf}
    \end{center}
    \caption{Vertices per second in UTS on \Grappa as the number of nodes grows.}
    \label{fig:grappa-uts}
\end{figure}

\TODO{UTS: story here is T1 class is bushy trees similar to properties
    of real world networks. Larger T1XXL (4.5 B vertices) provides
    more parallelism for continued scaling up. T3 class is deep skinny
    trees. Grappa can do no better than the straightforward UPC impl: there very little
    parallelism and the critical path is long}

@@

In order to put \Grappa's performance into a general context, we compare it
with XMT running BFS, PageRank, IntSort, GUPS and UTS. Since XMT is a
different hardware platform, we also compare \Grappa with hand-tuned MPI
versions of BFS and GUPS running on the same hardware. Finally, we also
compare it with UTS written for UPC. We run all experiments with 64 nodes.
